Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-05-15 AM+draft.tex
15 May added supmaterials

We analysed the simulated data with Eagle and the other computer programs, recording their memory usage and 
run (or elapse) times. The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM. Not all data generated under the five scenarios could be analysed by all implementations. Memory usage 
for many of the computer programs was the limiting factor (See Sup figure 1).  The single-locus program GEMMA was by 
far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
excess of the 128 Gbytes of available RAM for the analysis of data generated under 4000x1.5M and 10000x1.5M.  Even FaST-LMM, a single locus
implementation, required more than 128 Gbytes of memory for the analysis of the larger data sets 
when all the marker data was used to calculate the relationship matrix. Of the multiple-locus programs, only Eagle,  
with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
for data from scenario 10000x1.5M.

As the data sets became large, (4000x1.5m and 10000x1.5m), the multiple locus implementations memory footprint become excessive. Some even struggled with memory for data generated under 2000 x 500k. GEMMA, a single locus implementation had the lowest memory usage (sup figure 1). However, Eagle, with its ability to handle data larger than the memory capacity of the computer, was the only multiple locus implementation capable of producing findings for data from 10000x1.5m.

>>>>

We analysed the simulated data with Eagle and the other computer programs, recording their memory usage and 
run (or elapse) times. The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM. Not all data generated under the five scenarios could be analysed by all implementations. Memory usage 
for many of the computer programs was the limiting factor (See Sup figure 1).  The single-locus program GEMMA was by 
far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
excess of the 128 Gbytes of available RAM for the analysis of data generated under 4000x1.5M and 10000x1.5M.  Even FaST-LMM, a single locus
implementation, required more than 128 Gbytes of memory for the analysis of the larger data sets 
when all the marker data was used to calculate the relationship matrix. Of the multiple-locus programs, only Eagle,  
with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
for data from our largest scenario 10000x1.5M.
